\package{}: an \R{} package for implemetation of the ITHIM model

### 2016-09-16 manuscript/manuscript.tex
Updates to ITHIM manuscript 1

\section{Summary:}
\col[blue]{While increased active transportation, primarily cycling
  and walk, is often held up as a goal for improved public health —
  via reduced chronic disease, improved air quality, and reductions in
  greenhouse gas emissions --- obtaining estimates for the quantiative
  impacts of such shifts remains infrequent, in not altogether
  absent. Using U.S. metropolitan counties from the 10 regions defined
  by the Department of Health and Human Services (HHS1-10) we perform
  a comparative risk analysis between each regions and national
  averages as the baseline.  We investigated the reduced chronic
  disease burden of (x,y,z outcomes) from increased physical activity
  from population shifts in the amount of biking and walking doen for
  transport.  Our methods employ the Integrated Transport and Health
  Impacts Model (ITHIM) informed by regional travel parameters using
  the National Household Transportation Survey, non-travel activity
  metircs from the American Time Use Survey, and baseline Disease
  burden from the Centers for Disease Control and Prevention's WONDER
  database.  Our findings show that the West coast and the Northeast
  of the U.S. have travel patterns that, if mirrored by the nation,
  would save the U.S. thousands to hundreds of thousands DALYs.}

@@

The ITHIM model uses the comparative risk assesment framework to
estimate the population attributable fraction, $\af$.  In our case
$\af$ represents the proportional change in mortality and morbidity if
the physical activity levels were adjusted to represent an alternate
scenario.

@@

Here $R(x)$ represents the risk when the individual has exposure $x$,
in our case physical activity (\mets).  $P(x)$ and $Q(x)$ denote the
population density of individuals with exposure $x$, in the baseline
and alternate scenarios, respectively.  The exposure variable, $x$, is
modeled as the sum of travel-related exposure and non-travel-related
exposure.

@@

The time spent walking or cycling, i.e., active transport time, is
modeled as a log-normal random variable with means defined for each
age-sex stratum.  The non-travel-related physical activity is also
modeled as log-normal with stratum-specific means.  Estimates for
non-travel-related physical activity are diffiuclt.  We used the
American Time Use Survey data to estimate the relative means for each
of the age-sex strata.  We then treat the mean of the referent group
as a variable and compute results over a grid of values for this
variable.  See Figure \ref{dalyFigure}.

The travel-related exposure is estimated using active transport time
along with assumptions about how much physical activity is required
for cycling and walking.  Active transport time and non-travel-related
physical activity are modeled as independent random variables.  A
fixed value was used as the coefficient of variation for the
distributin of active transport time. A different, but also fixed,
value was used for the non-travel-related physical activity
distribution.

@@

\col[red]{\textit{
We chose to model road injuries in a more simplifed manner than the
authors of ITHIM.  The original ITHIM requires information about the
transportational infrastructure that cannot be accurately modeled on a
regional level, and is more well-suited to say a collection of
counties as a case study. \cite{maizlish2013}  We use the overall percent
of travel by walking and road fatalities per U.S. metropolitan area to
estimate road fatalaties in U.S. metropolitan areas given the percent
of travel by walking. \cite{UWTechReport}
}}

\col[red]{\textit{
We need to estimate the increase/decrease in road injuries/fatalaties
nation-wide as walking/cycling time increases.  We need a relative
risk of road fatalaty as a function of active transport time...
}}

\col[red]{\textit{

}}

@@

Can we model the decrease in \pEm{2.5} nationally as a function of
decrease in driving time?  The original plan was to use a baseline
value for \pEm{2.5} and a linear relationship bewteen change in
\pEm{2.5} and change in \vmt{}.

>>>>

\section{Summary:}
\col[blue]{While increased active transportation, primarily cycling
  and walk, is often held up as a goal for improved public health —
  via reduced chronic disease, improved air quality, and reductions in
  greenhouse gas emissions --- obtaining estimates for the quantiative
  impacts of such shifts remains infrequent, in not altogether
  absent. Using U.S. metropolitan counties from the 10 regions defined
  by the Department of Health and Human Services (HHS1-10) we perform
  a comparative risk analysis between each regions and national
  averages as the baseline.  We investigated the reduced chronic
  disease burden of ($x$, $y$, $z$ outcomes) from increased physical
  activity from population shifts in the amount of biking and walking
  doen for transport.  Our methods employ the Integrated Transport and
  Health Impacts Model (ITHIM) informed by regional travel parameters
  using the National Household Transportation Survey, non-travel
  activity metircs from the American Time Use Survey, and baseline
  Disease burden from the Centers for Disease Control and Prevention's
  WONDER database.  Our findings show that the West coast and the
  Northeast of the U.S. have travel patterns that, if mirrored by the
  nation, would save the U.S. thousands to hundreds of thousands
  DALYs.}

@@

\col[blue]{\textit{
    The ITHIM model uses the comparative risk
    assesment framework to estimate the population attributable
    fraction, $\af$.  In our case $\af$ represents the proportional
    change in mortality and morbidity if the physical activity levels
    were adjusted to represent an alternate scenario.
}}

@@

\col[blue]{\textit{
    Here $R(x)$ represents the risk when the individual has exposure $x$,
in our case physical activity (\mets).  $P(x)$ and $Q(x)$ denote the
population density of individuals with exposure $x$, in the baseline
and alternate scenarios, respectively.  The exposure variable, $x$, is
modeled as the sum of travel-related exposure and non-travel-related
exposure.
}}

@@

\col[blue]{\textit{
    The time spent walking or cycling, i.e., active transport time, is
modeled as a log-normal random variable with means defined for each
age-sex stratum.  The non-travel-related physical activity is also
modeled as log-normal with stratum-specific means.  Estimates for
non-travel-related physical activity are diffiuclt.  We used the
American Time Use Survey data to estimate the relative means for each
of the age-sex strata.  We then treat the mean of the referent group
as a variable and compute results over a grid of values for this
variable.  See Figure \ref{dalyFigure}.
}}

\col[blue]{\textit{
The travel-related exposure is estimated using active transport time
along with assumptions about how much physical activity is required
for cycling and walking.  Active transport time and non-travel-related
physical activity are modeled as independent random variables.  A
fixed value was used as the coefficient of variation for the
distributin of active transport time. A different, but also fixed,
value was used for the non-travel-related physical activity
distribution.
}}

@@

\col[blue]{\textit{
We chose to model road injuries in a more simplifed manner than the
authors of ITHIM.  The original ITHIM requires information about the
transportational infrastructure that cannot be accurately modeled on a
regional level, and is more well-suited to say a collection of
counties as a case study. \cite{maizlish2013}  We use the overall percent
of travel by walking and road fatalities per U.S. metropolitan area to
estimate road fatalaties in U.S. metropolitan areas given the percent
of travel by walking. \cite{UWTechReport}
}}

\col[blue]{\textit{
We need to estimate the increase/decrease in road injuries/fatalaties
nation-wide as walking/cycling time increases.  We need a relative
risk of road fatalaty as a function of active transport time...
}}

@@

\col[blue]{\textit{
    Can we model the decrease in \pEm{2.5} nationally as a function of
decrease in driving time?  The original plan was to use a baseline
value for \pEm{2.5} and a linear relationship bewteen change in
\pEm{2.5} and change in \vmt{}.
}}
